---
title: "The nasopharyngeal microbiota of elderly subjects with and without influenza-like illness"
subtitle: "Cohort Stats"
author: "Sudarshan A. Shetty"
date: "`r date()`"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---

This document describes the participant data.  

# Analysis  
**Load pkgs**    
```{r libraries, warning=FALSE, message=FALSE}
library(readr)
library(decontam)
#BiocManager::install("DECIPHER")
library(microbiome); 
#packageVersion("microbiome")
library(phyloseq)
#packageVersion("phyloseq")
library(dplyr)
#packageVersion("dplyr")
library(ggpubr)
#packageVersion("ggpubr")
library(patchwork)
#packageVersion("patchwork")
library(microbiomeutilities)
library(pheatmap)
library(factoextra)
library(janitor)

```

Create directory to store output from investigation of controls.   
```{r dir, eval=TRUE}
dir.create("03_cohort")
dir.create("03_cohort/figs")
dir.create("03_cohort/tabs")
#dir.create("tabs")
```

Color codes for each condition.  
```{r}
dis_col <- c(acute_ili="#6883ba", 
             control= "#166f65", 
             `NA`="black",
             acute_ili_14_days = "#6d597a", 
             control_14_days = "#87d2c9", 
             acute_ili_recovery = "#e56b6f",
             ili="steelblue") 

```


## read clean pseq 
```{r}

ps.main <- readRDS("data/rds/ps.clean.rds")

length(unique(sample_data(ps.main)$participant_id))

```

Three participants ELD156, ELD208, ELD39 have samples spread across multiple years as part atleast one ili and one control timepoint. These will will be removed from this paper. 

```{r , eval=TRUE}

ps.main <- subset_samples(ps.main, !(participant_id %in% c("ELD156", "ELD208", "ELD39")))

ps.main <- prune_taxa(taxa_sums(ps.main) >0, ps.main)

# Save this for rest of the anlaysis. 
saveRDS(ps.main, "data/rds/ps.clean.rds")

```


```{r}
print_ps(ps.main)
```


## Quick check on some variables  
```{r source-info }

knitr::kable(table(meta(ps.main)$source,meta(ps.main)$condition_group))
knitr::kable(table(meta(ps.main)$source,meta(ps.main)$moraxella_catarrhalis_2014))
```


Check per subject condition_status sampled. 
```{r subj-source}
metadf.so <- meta(ps.main) %>% 
   group_by(participant_id) %>% arrange(desc(condition_status)) %>% 
                    summarise(samples_for = paste(unique(condition_status), collapse = ","))
knitr::kable(table(metadf.so$samples_for))
```


Check for distribution of age and BMI 
```{r age-bmi-dist, fig.height=4, fig.width=6, fig.align="center"}
metadf <- meta(ps.main)
bmi_hist <- gghistogram(metadf, "BMI_2014", 
                        fill="condition_status", 
                        palette = dis_col, alpha=0.25,
                        add_density = TRUE)
age_hist <- gghistogram(metadf, "age_yrs_oct2014", 
                        fill="condition_status", 
                        palette = dis_col, alpha=0.25,
                        add_density = TRUE)

bmi_hist + age_hist + plot_annotation(tag_levels = "A") + 
  plot_layout(guides="collect") & theme(legend.position = 'top')  
ggsave("03_cohort/figs/bmi_age_hist.pdf", height = 4, width = 10)
```

```{r sex-condition-grp, fig.height=4, fig.width=4}

p <- plot_frequencies(sample_data(ps.main), "condition_group", "gender")
print(p + ggtitle("gender"))
ggsave("03_cohort/figs/gender.pdf", height = 4, width = 6)

knitr::kable(table(sample_data(ps.main)$condition_group, sample_data(ps.main)$gender ))

```

## Pick metadata  
```{r}

demo_data <- meta(ps.main)

# make a list of participants at different sampling times. 
# control
#total unique
length(unique(subset(demo_data, condition_group=="control")$participant_id))
ctrl.v1.participants <- subset(demo_data, condition_status=="control")$participant_id
ctrl.v2.participants <- subset(demo_data, condition_status=="control_14_days")$participant_id

# control
control.v1.df <- subset(demo_data, condition_status=="control")
control.v2.df <- subset(demo_data, condition_status=="control_14_days")

# now remove from v2 participants from v1
control.v2.df.s <- control.v2.df %>% 
  filter(!participant_id %in% ctrl.v1.participants)

control_df <- control.v1.df %>% bind_rows(control.v2.df.s)

# ILI 
#total unique
length(unique(subset(demo_data, condition_group=="ili")$participant_id))
ili.v1.participants <- subset(demo_data, condition_status=="acute_ili")$participant_id
ili.v2.participants <- subset(demo_data, condition_status=="acute_ili_14_days")$participant_id
ili.v3.participants <- subset(demo_data, condition_status=="acute_ili_recovery")$participant_id


ili.v1.df <- subset(demo_data, condition_status=="acute_ili")
ili.v2.df <- subset(demo_data, condition_status=="acute_ili_14_days")
ili.v3.df <- subset(demo_data, condition_status=="acute_ili_recovery")

# now remove from v2 participants from v1
ili.v2.df.s <- ili.v2.df %>% 
  filter(!participant_id %in% ili.v1.participants)

# now remove from v3 participants from v1
ili.v3.df.s <- ili.v3.df %>% 
  filter(!participant_id %in% ili.v1.participants)

# now remove from v3 participants from v2
ili.v3.df.s <- ili.v3.df.s %>% 
  filter(!participant_id %in% ili.v2.df.s$participant_id)
# Combine V1 V2 V3 ILI
iliv1v2v3 <-  bind_rows(ili.v1.df, ili.v2.df.s, ili.v3.df.s)

# Combine ILI and Control
demo_data <- bind_rows(iliv1v2v3,control_df)
write.csv(demo_data, "03_cohort/demo_data_cohort.csv")
saveRDS(demo_data, "03_cohort/demo_data_cohort.rds")
table(iliv1v2v3$moraxella_catarrhalis_2014, iliv1v2v3$condition_status)

```

```{r}

respiratory_disease <- demo_data %>% 
  group_by(condition_status, respiratory_disease_2014) %>% 
  tally() %>% 
  mutate(rel.freq = round(100 * n/sum(n), 1),
         rel.freq.pct = paste0(round(100 * n/sum(n), 0), "%"))
write_csv(respiratory_disease,"03_cohort/tabs/respiratory_disease.csv")


comorbidities <- demo_data %>% 
  group_by(condition_status, comorbidities_2014) %>% 
  tally() %>% 
  mutate(rel.freq = round(100 * n/sum(n), 1),
         rel.freq.pct = paste0(round(100 * n/sum(n), 0), "%"))
write_csv(comorbidities,"03_cohort/tabs/comorbidities.csv")
```

# Stats on metadata  

```{r}
library(coin)
library(xlsx)
```

## Create participant level data  

```{r}
#colnames(demo_data)
table(demo_data$condition_status)
# get unique participants
sub_part <- unique(demo_data$participant_id)

sub_data <- demo_data %>% distinct(participant_id, .keep_all = TRUE)

table(sub_data$condition_status)
dim(sub_data)
dim(demo_data)
```

```{r}
sub_data <- remove_empty(sub_data, which = c("rows","cols"))

```


# Check  

Identify variables with less or no difference  
```{r}
table(sub_data$transplant_2014, useNA = "always")
table(sub_data$hiv_infection_2014, useNA = "always")
table(sub_data$functional_asplenia_2014, useNA = "always")
table(sub_data$leukemia_2014, useNA = "always")
table(sub_data$lymphatic_cancer_2014, useNA = "always")
```

```{r}
sub_data <- sub_data %>% 
  dplyr::select(-c(transplant_2014, hiv_infection_2014,functional_asplenia_2014,leukemia_2014,lymphatic_cancer_2014))
```

# Group as ILI or No-ILI
```{r}
#sub_data$condition_group

good.index <- sub_data$participant_id[which(!is.na(sub_data$numb_of_symptoms))]
sub_data$ILI_ja_nee <- 0
sub_data$ILI_ja_nee[which(sub_data$participant_id %in% good.index)] <- 1
```


# Select variables of interest
```{r}
main_var <- c("ILI_ja_nee","gender", "age_yrs_oct2014")
comord.vars <- c("BMI_2014","respiratory_disease_2014", "comorbidities_2014",
                 "chronic_cvd_2014","Diabetes_2014", "autoimmune_disease_2014", 
                 "other_malignancies", "Roken_2014")

med.vars <- c("ace_inhibitors_2014", "statins_2014", "ppi_2014",
              "alpa_beta_blockers_2014","antimicrobials_2014", "arbs_2014", "insulin_2014",
              "Corticosteroiden_2014", "Antiepileptica_2014", "Antihistaminica_2014",
              "analgesics_2014", "vaccinated_2014", "flu_shot_2014_2015", "denzodiazepine_2014",
              "Antidepressiva_2014", "Glucoseverlagend_2014", "Stollingsremmers_2014", "Calciumregulatie_2014",
              "AstmaCOPD_2014", "number_of_medications_2014", "antibioitics_2014")
patho.vars <- colnames(sub_data)[134:155]


var.of.int <-  c(main_var,comord.vars,med.vars,patho.vars)
# create the final data set:
data.to.infer <- sub_data[, var.of.int]
```

## Age BMI stats

```{r}
# check age from No-ILI 
round(mean(subset(data.to.infer, ILI_ja_nee == 0)$age_yrs_oct2014),1)
round(sd(subset(data.to.infer, ILI_ja_nee == 0)$age_yrs_oct2014),1)

# check age from ILI 
round(mean(subset(data.to.infer, ILI_ja_nee == 1)$age_yrs_oct2014),1)
round(sd(subset(data.to.infer, ILI_ja_nee == 1)$age_yrs_oct2014),1)

# check bmi from No-ILI 
round(mean(subset(data.to.infer, ILI_ja_nee == 0)$BMI_2014),1)
round(sd(subset(data.to.infer, ILI_ja_nee == 0)$BMI_2014),1)

# check bmi from ILI 
round(mean(subset(data.to.infer, ILI_ja_nee == 1)$BMI_2014, na.rm = T),1)
round(sd(subset(data.to.infer, ILI_ja_nee == 1)$BMI_2014,na.rm = T),1)

# check gender  
table(subset(data.to.infer, ILI_ja_nee == 0)$gender)
table(subset(data.to.infer, ILI_ja_nee == 1)$gender)
```


```{r eval=FALSE}
##################
# overview of new variables
table(sub_data$antimicrobials_2014, sub_data$ILI_ja_nee, useNA = "always")
table(sub_data$antimicrobials_2014, sub_data$antimicrobial_during_sampling_2014, useNA = "always")
table(data.to.infer$antimicrobial_during_sampling_2014, data.to.infer$ILI_ja_nee, useNA = "always")
table(data.to.infer$antibioitics_2014, data.to.infer$ILI_ja_nee, useNA = "always")

```

```{r}
###################
# Create age.group:
range(data.to.infer$age_yrs_oct2014)
age.breaks <- seq(60,90,5)
data.to.infer <- base::transform(data.to.infer,
                           age.group=cut(age_yrs_oct2014, 
                                         breaks=age.breaks, 
                                         right=TRUE, 
                                         include.lowest=TRUE))
data.to.infer$age.group <- as.character(data.to.infer$age.group); table(data.to.infer$age.group)
#
```

```{r}
#################################
# Create strata of "confounders":
#names(data.to.infer)[2] <- "gender"
confounders <- c("gender","age.group")
name.of.stratification <- "gender-age-group"

aux.data.set <- subset(data.to.infer,select="age.group")
aux.data.set <- base::transform(aux.data.set,stratum=rep(NA,nrow(aux.data.set)))
for(i in (1:nrow(aux.data.set))){
  aux.data.set[i,]$stratum <- paste(aux.data.set[i,-length(aux.data.set)],collapse="/")
}
aux.data.set$stratum <- as.character(aux.data.set$stratum)
# str(aux.data.set); head(aux.data.set)
#
data.to.infer <- base::transform(data.to.infer,stratum=aux.data.set$stratum)
data.to.infer$stratum <- as.character(data.to.infer$stratum); table(data.to.infer$stratum)
# names(data.set); str(data.set); head(data.set)

```

```{r}
frequency.of.strata <- table(data.to.infer$stratum); mean(frequency.of.strata); frequency.of.strata
#
data.to.infer$stratum <- as.factor(data.to.infer$stratum)

nominal.bound.on.FDR <- 0.2
```

```{r}

colnames(data.to.infer)[1:10]
set.seed(1977)
B <- 100000
#
tests <- data.frame(var.1 = c("gender",
                              "age_yrs_oct2014",
                              "BMI_2014"),
                    var.2 = c("ILI_ja_nee", "ILI_ja_nee", "ILI_ja_nee"),
                    p.value = c(as.numeric(fisher.test(data.to.infer$gender,                                                       as.factor(data.to.infer$ILI_ja_nee))$p.value),
                                as.numeric(t.test(age_yrs_oct2014~ILI_ja_nee, 
                                                  data =  data.to.infer)$p.value),
                                as.numeric(pvalue(independence_test(BMI_2014~ ILI_ja_nee, data=data.to.infer,distribution=approximate(B))))),
                    test = c("Fisher","T test", "Sum statistic-no blocking by stratum"))

tests_1 <- tests

#
# test separately information related to antibiotica use:
table(data.to.infer$antibioitics_2014, data.to.infer$ILI_ja_nee, useNA = "always")
table(data.to.infer$antimicrobials_2014, data.to.infer$ILI_ja_nee, useNA = "always")
response <- data.to.infer[,"antimicrobials_2014"]
CHM.test <- cmh_test(as.factor(response)~as.factor(ILI_ja_nee)|stratum, data=data.to.infer,distribution=approximate(B))
p.value <- as.numeric(pvalue(CHM.test))

tests <- data.frame(var.1 = c("antibioitics_2014","antimicrobials_2014"),
                       var.2 = c("ILI_ja_nee", "ILI_ja_nee"),
                       p.value = c(as.numeric(pvalue(cmh_test  (as.factor(data.to.infer[,"antibioitics_2014"])~as.factor(ILI_ja_nee)|stratum, data=data.to.infer,distribution=approximate(B)))),                                   as.numeric(pvalue(cmh_test(as.factor(data.to.infer[,"antimicrobials_2014"])~as.factor(ILI_ja_nee)|stratum, data=data.to.infer,distribution=approximate(B))))),
                       test = c("Cochran-Mantel-Haenszel","Cochran-Mantel-Haenszel"))
print(tests)
```



```{r}
#transplant_2014 (11) number_cigarettes_week_2014 (19) antimicrobial_during_sampling_2014(27)
data.to.infer.ori <- data.to.infer
data.to.infer.ori <- remove_constant(data.to.infer.ori)

#data.to.infer.ori <- data.to.infer
#colnames(data.to.infer.ori)
#colnames(data.to.infer)
data.to.infer <- data.to.infer.ori %>% dplyr::select(-all_of(patho.vars))
table(data.to.infer.ori$duration_symptom_before_collection)
#data.to.infer <- data.to.infer.ori
#data.to.infer <- data.to.infer[,-c(12,)] #,51,52,53,
tests <- NULL 
for (i in 5:(ncol(data.to.infer)-3)) {
  response <- data.to.infer[,i]
  CHM.test <- cmh_test(as.factor(response)~as.factor(ILI_ja_nee)|stratum, data=data.to.infer,distribution=approximate(B))
  p.value <- as.numeric(pvalue(CHM.test))
  
  output.p <- data.frame(var.1=paste0(names(data.to.infer)[i]),
                         var.2="ILI_ja_nee",
                         p.value=p.value,
                         test="Cochran-Mantel-Haenszel")
  print(output.p)
  tests <- rbind(tests,output.p)      
}
tests 
```

```{r}
multiple.tests <- tests
multiple.tests <- multiple.tests[order(multiple.tests$p.value),]; row.names(multiple.tests) <- NULL
aux.histogram <- hist(multiple.tests$p.value,plot=T)
gamma.hat <- min(1,aux.histogram$density[length(aux.histogram$density)]); gamma.hat
multiple.tests <- base::transform(multiple.tests,rank=(1:nrow(multiple.tests)))
bound.FDR <- nrow(multiple.tests)*multiple.tests$p.value/(1:nrow(multiple.tests))
better.bound.FDR <- gamma.hat*nrow(multiple.tests)*multiple.tests$p.value/(1:nrow(multiple.tests))
multiple.tests <- base::transform(multiple.tests,
                            bound.FDR=bound.FDR,
                            better.bound.FDR=better.bound.FDR)
multiple.tests
multiple.tests.medictions <- multiple.tests
```

```{r}
file.name <- paste("03_cohort/tabs/associations.between.medication.and.ILI.status.txt",sep=".")
write.table(multiple.tests,file=file.name,row.names=FALSE,sep=";")

file.name <- paste("03_cohort/tabs/associations.between.medication.and.ILI.status.xlsx",sep=".")
write.xlsx(multiple.tests,file=file.name,row.names=FALSE)

```

 
## pathogens  
```{r}
data.to.infer.vir <- data.to.infer.ori[, c(main_var,patho.vars,"stratum")]
data.to.infer.vir <- replace(data.to.infer.vir, data.to.infer.vir == "unknown", 2)
data.to.infer.vir <- replace(data.to.infer.vir, data.to.infer.vir ==  3, 2)

#summary(data.to.infer.vir)
data.to.infer.vir <- remove_constant(data.to.infer.vir)

data.to.infer.vir <- data.to.infer.vir[,-c(2,3,14,17)] #14 influenza_type_2014
colnames(data.to.infer.vir)
table(data.to.infer.vir$rhinovirus_2014)
#data.to.infer.vir <- data.to.infer.vir[,-c(1:3,25,20)]

test_vir <- c()
for (i in 2:(ncol(data.to.infer.vir)-1)) {
  response <- data.to.infer.vir[,i]
  print(colnames(data.to.infer.vir)[i])
  CHM.test <- cmh_test(as.factor(response) ~ as.factor(ILI_ja_nee)|stratum, 
                       data=data.to.infer.vir, 
                       distribution=approximate(B))
  
  p.value <- as.numeric(pvalue(CHM.test))
  
  output.p <- data.frame(var.1=paste0(names(data.to.infer.vir)[i]),
                         var.2="ILI_ja_nee",
                         p.value=p.value,
                         test="Cochran-Mantel-Haenszel")
  print(output.p)
  test_vir <- rbind(test_vir,output.p)      
}

multiple.tests <- test_vir
multiple.tests <- multiple.tests[order(multiple.tests$p.value),]; row.names(multiple.tests) <- NULL
aux.histogram <- hist(multiple.tests$p.value,plot=T)
gamma.hat <- min(1,aux.histogram$density[length(aux.histogram$density)]); gamma.hat
multiple.tests <- base::transform(multiple.tests,rank=(1:nrow(multiple.tests)))
bound.FDR <- nrow(multiple.tests)*multiple.tests$p.value/(1:nrow(multiple.tests))
better.bound.FDR <- gamma.hat*nrow(multiple.tests)*multiple.tests$p.value/(1:nrow(multiple.tests))
multiple.tests <- base::transform(multiple.tests,
                            bound.FDR=bound.FDR,
                            better.bound.FDR=better.bound.FDR)
multiple.tests

file.name <- paste("03_cohort/tabs/associations.between.pathogens.and.ILI.status.txt",sep=".")
write.table(multiple.tests,file=file.name,row.names=FALSE,sep=";")

file.name <- paste("03_cohort/tabs/associations.between.pathogens.and.ILI.status.xlsx",sep=".")
write.xlsx(multiple.tests,file=file.name,row.names=FALSE)

```


```{r eval=FALSE}
data.to.infer.numbs <- data.to.infer.vir[,c(1:3,26,27,28)]

#test for number of virus seperately
#"numb_pathogens" "numb_viruses"   "numb_bacteria"  "ILI_ja_nee"     "age.group"      "stratum" 

data.to.infer.numbs <- replace(data.to.infer.numbs, data.to.infer.numbs == "unknown", 0)

var1.numb_pathogens <- cmh_test(as.factor(data.to.infer.numbs[,1])~as.factor(ILI_ja_nee)|stratum, data=data.to.infer.numbs,distribution=approximate(B))
var1.numb_viruses <- cmh_test(as.factor(data.to.infer.numbs[,2])~as.factor(ILI_ja_nee)|stratum, data=data.to.infer.numbs,distribution=approximate(B))
var1.numb_bacteria <- cmh_test(as.factor(data.to.infer.numbs[,3])~as.factor(ILI_ja_nee)|stratum, data=data.to.infer.numbs,distribution=approximate(B))


```

### Medications summary  
```{r}

library(arsenal)
table_one <- tableby(ILI_ja_nee ~ ., data = data.to.infer)
#head(table_one)

write2word(table_one, "~/respiratory/RespiratoryMicrobiotaElderly/03_cohort/arsenal_table.doc",
  keep.md = TRUE,
  quiet = TRUE, # passed to rmarkdown::render
  title = "Summary from data.to.infer") # passed to summary.tableby


```


### pathogens summary  
```{r}

table_two <- tableby(ILI_ja_nee ~ ., data = data.to.infer.vir)
#head(table_one)

write2word(table_two, "~/respiratory/RespiratoryMicrobiotaElderly/03_cohort/arsenal_table_pathogens.doc",
  keep.md = TRUE,
  quiet = TRUE, # passed to rmarkdown::render
  title = "Summary from data.to.infer.vir") # passed to summary.tableby


```


## Table of baseline   
```{r}
table(data.to.infer$comorbidities_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Autoimmune disease
table(data.to.infer$autoimmune_disease_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Cardiovascular disease
table(data.to.infer$chronic_cvd_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Diabetes
table(data.to.infer$Diabetes_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Malignancies
table(data.to.infer$other_malignancies, data.to.infer$ILI_ja_nee, useNA = "always")

#Renal disease
#table(data.to.infer$chronic_kidney_disease_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Respiratory disease
table(data.to.infer$respiratory_disease_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Smoking
table(data.to.infer$Roken_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Influenza vaccination
table(data.to.infer$flu_shot_2014_2015, data.to.infer$ILI_ja_nee, useNA = "always")
#ACE inhibitors
table(data.to.infer$ace_inhibitors_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Alpha-/betablocker
table(data.to.infer$alpa_beta_blockers_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Analgesic
table(data.to.infer$analgesics_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#Antidepressants
table(data.to.infer$Antidepressiva_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#anticoagulant
table(data.to.infer$Stollingsremmers_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Antiepileptic
table(data.to.infer$Antiepileptica_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Antihistamines
table(data.to.infer$Antihistaminica_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#ARBs
table(data.to.infer$arbs_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Asthma/COPD medication
table(data.to.infer$AstmaCOPD_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Calcium medication
table(data.to.infer$Calciumregulatie_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Corticosteroids
table(data.to.infer$Corticosteroiden_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Diuretics
#Dopaminergic
#table(data.to.infer$Dopaminergica_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Gonadocorticoids
#table(data.to.infer$Geslachtshormonen_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Immunosuppressants
#table(data.to.infer$immunosuppressants_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Insulin
table(data.to.infer$insulin_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Other cancer medication
#table(data.to.infer$cancer_medication_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#PPIs
table(data.to.infer$ppi_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#Statins
table(data.to.infer$statins_2014, data.to.infer$ILI_ja_nee, useNA = "always")

#T2 diabetes medication
table(data.to.infer$number_of_medications_2014, data.to.infer$ILI_ja_nee, useNA = "always")
#denzodiazepine
table(data.to.infer$denzodiazepine_2014, data.to.infer$ILI_ja_nee, useNA = "always")

```

## Table of pathogens/virus  

```{r}
#No. of pathogens detected
#table(data.to.infer.ori$numb_pathogens, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Influenza virus
table(data.to.infer.ori$influenza_2014_any, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Influenza A virus
table(data.to.infer.ori$influenza_A_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Influenza B virus
table(data.to.infer.ori$influenza_B_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")

#table(data.to.infer.ori$influenza_type_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Rhinovirus
table(data.to.infer.ori$rhinovirus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Coronavirus
table(data.to.infer.ori$coronavirus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#RSV
table(data.to.infer.ori$respiratory_syncytial_virus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#HPIV
table(data.to.infer.ori$human_parainfluenzavirus_2014_any, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#HMPV
table(data.to.infer.ori$human_metapneumovirus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Mycoplasma pneumoniae
table(data.to.infer.ori$mycoplasma_pneumoniae_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#Chlamydophila pneumoniae
table(data.to.infer.ori$chlamydophila_pneumoniae_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~Haemophilus
table(data.to.infer.ori$haemophilus_2014_any, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~H. influenzae
table(data.to.infer.ori$haemophilus_influenzae_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~H. haemolyticus
table(data.to.infer.ori$haemophilus_haemolyticus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~Haemolytic Streptococcus
table(data.to.infer.ori$haemolytic_streptococcus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~Staphylococcus aureus
table(data.to.infer.ori$staphylococcus_aureus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~Moraxella catarrhalis
table(data.to.infer.ori$staphylococcus_aureus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
#~Streptococcus pneumoniae
table(data.to.infer.ori$streptococcus_pneumoniae_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")

#adeno_2014
#bocavirus_2014
#bordetella_pertussis_2014
#legionella_pneumophila_2014
table(data.to.infer.ori$legionella_pneumophila_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
table(data.to.infer.ori$bordetella_pertussis_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
table(data.to.infer.ori$bocavirus_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
table(data.to.infer.ori$adeno_2014, data.to.infer.ori$ILI_ja_nee, useNA = "always")
```


```{r}
sessionInfo()
```



