---
title: "The nasopharyngeal microbiota of elderly subjects with and without influenza-like illness"
subtitle: "Microbiota stability analysis"
author: "Sudarshan A. Shetty"
date: "`r date()`"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
    toc_float: true
---

# Microbiota stability 
# Analysis   

## Load pkgs    
```{r libraries, warning=FALSE, message=FALSE}

library(microbiome); #packageVersion("microbiome")
library(phyloseq)#packageVersion("phyloseq")
library(dplyr)#packageVersion("dplyr")
library(ggpubr)#packageVersion("ggpubr")
library(patchwork)#packageVersion("patchwork")
library(vegan)
library(gghalves)
library(reshape)
library(microbiomeutilities)
library(MiSPU)
library(rstatix)
library(ggridges)
library(tibble)
theme_set(theme_bw())
source("codes/generalCodes.R")
source("codes/coreAnalysis.R")
dir.create("05_stability")

```


Read the phyloseq object created `01_ili_deomographics.rmd`    

```{r var-cols}

dis_col <- c(acute_ili="#6883ba", control = "#166f65", `NA`="black",
             acute_ili_14_days = "#6d597a", control_14_days = "#87d2c9", 
             acute_ili_recovery = "#e56b6f") 

```

## Import  

```{r}
ps.main <- readRDS("data/rds/ps.clean.rds")

length(unique(sample_data(ps.main)$participant_id))

#any(unique(sample_data(ps.main)$participant_id %in% c("ELD156", "ELD208", "ELD39")))
```

397 participants. 

Next, we compare stability of microbiota composition in non-ILI group with that of ILI group.  

## Root tree  
```{r}
set.seed(12580)
my.tree <- phy_tree(ps.main)
out.group <- pick_new_outgroup(my.tree)
out.group ## [1] "ASV2239" Genus:"Beijerinckiaceae"
out.group <- "ASV8595"
tax_table(ps.main)[out.group,]
new.tree <- ape::root(my.tree, outgroup=out.group, resolve.root=TRUE)
phy_tree(ps.main) <- new.tree
phy_tree(ps.main)

```



# Microbiota stability   

```{r}
#min(sample_sums(ps.main))=1975
set.seed(18474)
ps.main.rar <- rarefy_even_depth(ps.main, sample.size = min(sample_sums(ps.main)))
#408OTUs were removed because they are no longer
```


## GUniFrac  

Use rarefied data

```{r}
# Use GUniFrac from MiSPU R pks as it is faster. 
# GUniFrac
gunifracs.stab <- GUniFrac(t(abundances(ps.main.rar)), 
                      tree=ps.main.rar@phy_tree, 
                      alpha=c(0, 0.5, 1))

saveRDS(gunifracs.stab, "data/rds/gunifracs.stab.rds")

guni_weighted_0.5 <- gunifracs.stab$d5
colnames(guni_weighted_0.5) <- sample_names(ps.main.rar)
rownames(guni_weighted_0.5) <- sample_names(ps.main.rar)

```

### Prep table   
```{r}

# get metadata 
meta_tib <- get_tibble(ps.main.rar, slot="sam_data") %>% 
   dplyr::select(participant_id, column_id, 
                 condition_status, condition_group, 
                 gender, age_yrs_oct2014)
table(meta_tib$condition_group)
```


## Pick participants with samples  
### Control visit 1 and 2  

```{r}
#ps.main.gen <- aggregate_taxa(ps.main, "genus")
ps.no.ili <- subset_samples(ps.main, condition_group=="control")
no.ili.samples <- meta(subset_samples(ps.no.ili, condition_status == "control"))$participant_id
no.ili.v2.samples <- meta(subset_samples(ps.no.ili, condition_status == "control_14_days"))$participant_id
# present in both
no.ili.v1.v2 <- intersect(no.ili.samples, no.ili.v2.samples)

ps.no.ili.v1v2 <-  subset_samples(ps.no.ili, participant_id %in% no.ili.v1.v2)
ps.no.ili.v1v2 <- prune_taxa(taxa_sums(ps.no.ili.v1v2) > 0 , ps.no.ili.v1v2)
```


### Dist table Control visit 1 and 2  

```{r}

#ctrl_gunifrac <- guni_weighted_0.5[]

cntrl_v1v2 <- reshape2::melt(
  as.matrix(guni_weighted_0.5), 
  varnames = c("S1", "S2"),
  value.name = "GUniFrac") %>%
  #filter(S1 != S2) %>% 
  mutate_if(is.factor, as.character) %>% 
  left_join(meta_tib, by = c("S1" = "column_id")) %>%
  left_join(meta_tib, by = c("S2" = "column_id"),suffix = c("_1", "_2")) %>% 
  filter(S1 != S2) %>% 
  filter(participant_id_1 %in% no.ili.v1.v2) %>% 
  filter(participant_id_2 %in% no.ili.v1.v2) %>% 
  filter(participant_id_1 == participant_id_2) %>%
  mutate(SubjectID = participant_id_1) %>%
  distinct(SubjectID, .keep_all = TRUE) %>% 
  dplyr::select(-participant_id_1, -participant_id_1) %>% 
  mutate(stability = 1-GUniFrac, comparison = "Control v1 vs v2")

```

### ILI visit 1 and 2   

We make a list of participants, then find those in all three visits and then make phyloseq object with ili participants with all three timepoints. the phyloseq objects are for later analysis.  
```{r}
# get subject who have acute ssamples
ps.ili <- subset_samples(ps.main, condition_group=="ili")

# list subject with acute samples
acute_samples <- microbiome::meta(subset_samples(ps.ili, 
                                                 condition_status == "acute_ili"))$participant_id
# list subject with acute 14d samples
ili_v2_samples <- microbiome::meta(subset_samples(ps.ili, 
                                                  condition_status == "acute_ili_14_days"))$participant_id
# list subject with acute recovery samples
ili_v3_samples <- microbiome::meta(subset_samples(ps.ili, 
                                                  condition_status == "acute_ili_recovery"))$participant_id
# present in both
ili_v1_v2 <- intersect(acute_samples, ili_v2_samples)
ili_v1_v2_v3 <- intersect(ili_v1_v2, ili_v3_samples)
# keep only those subject with illi 1,2 3 samples
ps.ili.v1v2v3 <- subset_samples(ps.ili, participant_id %in% ili_v1_v2_v3)
# 141075010361 is baseline from vaccine group 
#table(meta(ps.ili.v1v2v3)$participant_id)

# ili v1 v2
ps.ili.v12 <- subset_samples(ps.ili.v1v2v3, condition_status != "acute_ili_recovery")
ps.ili.v12 <- prune_taxa(taxa_sums(ps.ili.v12) > 0 , ps.ili.v12)
# ili v2 v3
ps.ili.v23 <- subset_samples(ps.ili.v1v2v3, condition_status != "acute_ili")
ps.ili.v23 <- prune_taxa(taxa_sums(ps.ili.v23) > 0 , ps.ili.v23)
# ili v1 v3
ps.ili.v13 <- subset_samples(ps.ili.v1v2v3, condition_status != "acute_ili_14_days")
ps.ili.v13 <- prune_taxa(taxa_sums(ps.ili.v13) > 0 , ps.ili.v13)

```

### Dist table ILI visit 1 and 2  
```{r}

ili_v1v2 <- reshape2::melt(
  as.matrix(guni_weighted_0.5), 
  varnames = c("S1", "S2"),
  value.name = "GUniFrac") %>%
  #filter(S1 != S2) %>% 
  mutate_if(is.factor, as.character) %>% 
  left_join(meta_tib, by = c("S1" = "column_id")) %>%
  left_join(meta_tib, by = c("S2" = "column_id"),suffix = c("_1", "_2")) %>% 
  filter(S1 != S2) %>% 
  filter(S1 %in% sample_names(ps.ili.v12)) %>% 
  filter(S2 %in% sample_names(ps.ili.v12)) %>% 
  filter(participant_id_1 == participant_id_2) %>%
  mutate(SubjectID = participant_id_1) %>%
  distinct(SubjectID, .keep_all = TRUE) %>% 
  dplyr::select(-participant_id_1, -participant_id_1) %>% 
  mutate(stability = 1-GUniFrac, comparison = "ILI v1 vs v2")
```

### Dist table ILI visit 2 and 3  
```{r}

ili_v2v3 <- reshape2::melt(
  as.matrix(guni_weighted_0.5), 
  varnames = c("S1", "S2"),
  value.name = "GUniFrac") %>%
  #filter(S1 != S2) %>% 
  mutate_if(is.factor, as.character) %>% 
  left_join(meta_tib, by = c("S1" = "column_id")) %>%
  left_join(meta_tib, by = c("S2" = "column_id"),suffix = c("_1", "_2")) %>% 
  filter(S1 != S2) %>% 
  filter(S1 %in% sample_names(ps.ili.v23)) %>% 
  filter(S2 %in% sample_names(ps.ili.v23)) %>% 
  filter(participant_id_1 == participant_id_2) %>%
  mutate(SubjectID = participant_id_1) %>%
  distinct(SubjectID, .keep_all = TRUE) %>% 
  dplyr::select(-participant_id_1, -participant_id_1) %>% 
  mutate(stability = 1-GUniFrac, comparison = "ILI v2 vs v3")
```

### Dist table ILI visit 1 and 3  
```{r}

ili_v1v3 <- reshape2::melt(
  as.matrix(guni_weighted_0.5), 
  varnames = c("S1", "S2"),
  value.name = "GUniFrac") %>%
  #filter(S1 != S2) %>% 
  mutate_if(is.factor, as.character) %>% 
  left_join(meta_tib, by = c("S1" = "column_id")) %>%
  left_join(meta_tib, by = c("S2" = "column_id"),suffix = c("_1", "_2")) %>% 
  filter(S1 != S2) %>% 
  filter(S1 %in% sample_names(ps.ili.v13)) %>% 
  filter(S2 %in% sample_names(ps.ili.v13)) %>% 
  filter(participant_id_1 == participant_id_2) %>%
  mutate(SubjectID = participant_id_1) %>%
  distinct(SubjectID, .keep_all = TRUE) %>% 
  dplyr::select(-participant_id_1, -participant_id_1) %>% 
  mutate(stability = 1-GUniFrac, comparison = "ILI v1 vs v3")
```

### Merge stability tables  
```{r}
gunifrac_stability <- cntrl_v1v2 %>% 
  bind_rows(ili_v1v2) %>% 
  bind_rows(ili_v2v3) %>% 
  bind_rows(ili_v1v3)

gunifrac_stability$comparison <- factor(gunifrac_stability$comparison, 
                                        levels = c("Control v1 vs v2", 
                                                   "ILI v1 vs v2", 
                                                   "ILI v2 vs v3", 
                                                   "ILI v1 vs v3"))

```


### Plot stability  

## Plot stability  

```{r}

# make a pairwise list that we want to compare.
comp_pairs <- make_pairs(gunifrac_stability$comparison)

comp_cols <- c(`Control v1 vs v2` = "#9e0059",
               `ILI v1 vs v2` = "#b56576",
               `ILI v2 vs v3` = "#eaac8b",
               `ILI v1 vs v3` = "#7a87bf")

#ggdensity(v123, "stability", facet.by = "condition_group_1")
```


```{r}

stab.plot <- ggplot(gunifrac_stability,
                    aes(y = stability, x = comparison, 
                        fill = comparison)) + 
  geom_boxplot(aes(fill = comparison), width = 0.2, 
               outlier.shape = NA, alpha = 0.25) + 
  geom_jitter(aes(group = comparison, color = comparison), 
              alpha = 0.25, size = 1) + 
  geom_half_violin(position = position_nudge(x = 0.15, y = 0), 
                   alpha = 0.2, side = "r") + 
  #stat_compare_means(comparisons = comp_pairs, 
   #                  hide.ns = TRUE) + 
  scale_fill_manual("Comparisons",values=comp_cols) +
  scale_color_manual("Comparisons",values=comp_cols) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("") + ylab("Stability (1-GUniFrac)") + 
  theme_biome_utils() + 
  theme(#legend.position = "top",
    text = element_text(size=14)) 

stab.plot
ggsave("05_stability/stability_gunifrac.pdf", height = 5, width = 5)
ggsave("05_stability/stability_gunifrac.tiff", height = 5, width = 5)
```

## Analayse stability groups  

### Groups/bins  
#### BC
```{r}
#http://web.mnstate.edu/peil/MDEV102/U4/S36/S363.html
# I use sumary stats 
gunifrac_stability_groups <- gunifrac_stability %>% 
  mutate(stability_group = ifelse(stability >= summary(gunifrac_stability$stability)[5], "High", 
                                  ifelse(stability <= summary(gunifrac_stability$stability)[2], "Low", "Medium")))

write.table(gunifrac_stability_groups, "05_stability/ili_3_microbiota_stability.txt", sep = "\t")

#Count stability subjects
bc_stab_ov <- gunifrac_stability_groups %>% 
  #distinct(participant_id, .keep_all = TRUE) %>% 
  group_by(comparison,stability_group) %>% tally() %>% 
  mutate(rel.freq = round(100 * n/sum(n), 1),
         rel.freq.pct = paste0(round(100 * n/sum(n), 0), "%")) 
knitr::kable(bc_stab_ov)

bc_stab_ov_wd <- bc_stab_ov %>% 
  tidyr::pivot_wider(id_cols=stability_group, names_from = comparison, values_from="rel.freq")
bc_stab_ov_wd$Measure <- "GUniFrac"
```


## Intra-individual dynamics in stability   

```{r}
ps.no.ili.v1v2@phy_tree <- NULL
ps.ili.v12@phy_tree <- NULL
ps.ili.v23@phy_tree <- NULL
ps.ili.v13@phy_tree <- NULL

ps.merge <- merge_phyloseq(ps.no.ili.v1v2,ps.ili.v12, ps.ili.v23,ps.ili.v13)
ps.merge <- prune_taxa(taxa_sums(ps.merge) >0, ps.merge)

# re-add tree from main source as diff tips in each object.
phy_tree(ps.merge) <- phy_tree(ps.main)
length(unique(sample_data(ps.merge)$participant_id))
saveRDS(ps.merge, "data/rds/ps.merge.rds")

```


```{r}
#make data smaller with select col

# order groups
gunifrac_stability_groups$stability_group <- factor(gunifrac_stability_groups$stability_group,
                                                    levels = c("High", "Medium","Low"))

```

### Intra-and inter- individual variation  
```{r}

cols_m <- c(High="#247ba0", Medium = "#0b2545", Low="#db3a34")
stab_pairs <- make_pairs(as.character(gunifrac_stability_groups$comparison))

stab.line.plot <- gunifrac_stability_groups %>% 
  filter(comparison != "Control v1 vs v2") %>% 
  #mutate(Group=ifelse(condition_group_1=="no_ili_v1_v2", "NO-ILI", "ILI")) %>% 
  ggplot(aes(comparison, stability)) + 
  geom_line(aes(group=SubjectID), color="grey", alpha=0.5) + 
  geom_jitter(aes(color=stability_group), 
              alpha=0.25, size=2, width = 0.05) + 
  ylab("Stability (1-GUniFrac)") +
  #facet_grid(~Group) + 
  #stat_compare_means(comparisons = , paired = TRUE) +
  scale_color_manual("Stability",values = cols_m) +
  theme_biome_utils() + theme(legend.position = "top")

ggsave("05_stability/03_Individua_combi_stability.pdf", height = 4, width = 7)
ggsave("05_stability/03_Individua_combi_stability.tiff", height = 4, width = 7)

```

## Alpha diversity and stability   

```{r}
ps.main.rar <- phyloDiversity(ps.main.rar)
dom.rar <- core_abundance(ps.main.rar, 0.00001, 0.75)
sample_data(ps.main.rar)$core_dom <- dom.rar 
stab.samples <- union(gunifrac_stability_groups$S1,gunifrac_stability_groups$S2)

ps.main.rar.2 <- prune_samples(stab.samples, ps.main.rar)
ps.stab.v1.v2 <- subset_samples(ps.main.rar.2, condition_status %in% c("control", "acute_ili"))
table(meta(ps.stab.v1.v2)$condition_status)
length(unique(meta(ps.stab.v1.v2)$participant_id))
```

```{r}
ps.stab.meta <- meta(ps.stab.v1.v2) 
# pick v1 v2 comparisons
stab.cb.v1 <- gunifrac_stability_groups %>% 
  #filter(condition_group_1 %in% c("no_ili_v1_v2","ili_v1_v2")) %>% 
  left_join(ps.stab.meta, by=c("SubjectID"="participant_id"))
```


### Plot diversity vs stability   

```{r fig.width=12, fig.height=4}
p.dv <- ggplot(stab.cb.v1, aes(stability,PD)) +
  geom_point(aes(color=comparison), alpha=0.5, size=3) +
  facet_grid(~comparison) +
  scale_color_manual(values=comp_cols) +
  theme_biome_utils() +
  stat_cor(method = "spearman", label.y = 40) +
  geom_smooth(method = "lm") +
  theme(legend.position = "none", strip.text = element_text(size=14)) + 
  ylab("Phylogenetic diversity") + xlab("Stability (1-GUniFrac)") + rotate_x_text()
p.dv
ggsave("05_stability/stability.spearman.cor.pdf", height = 3, width = 7)
```

```{r}
p.dv.dom <- ggplot(stab.cb.v1, aes(stability,core_dom)) +
  geom_point(aes(color=comparison), alpha=0.5, size=2) +
  facet_grid(~comparison) +
  scale_color_manual(values=comp_cols) +
  theme_minimal() +
  stat_cor(method = "spearman", label.y = 1.5) +
  geom_smooth(method = "lm") +
  theme(legend.position = "none", strip.text = element_text(size=14)) + 
  ylab("Abundance \ncore microbiota") + xlab("Stability (1-GUniFrac)") + 
  rotate_x_text()
p.dv.dom 
ggsave("msfigs/core_stability.bet.alpha.spearman.cor.pdf", height = 4, width = 8)
```

#### Figure 
```{r}
stab.plot.f <-  stab.plot + theme_minimal(base_size = 10) + theme(legend.position="none")
stab.line.plot.f <- stab.line.plot + theme_minimal(base_size = 10)
p.dv.f <- p.dv + theme_minimal(base_size = 10)
p.dv.dom <- p.dv.dom + theme_minimal(base_size = 10) + theme(legend.position="none")
(stab.plot.f+stab.line.plot.f)/p.dv.f /p.dv.dom + plot_annotation(tag_levels = "A")
ggsave("msfigs/stability.bet.alpha.spearman.cor.pdf", height = 8, width = 10)
```


```{r}
compar.stbs <- make_pairs(stab.cb.v1$stability_group)
head(stab.cb.v1)
ggplot(stab.cb.v1, aes(stability_group,PD)) +
  geom_boxplot() + 
  facet_grid(~comparison) +
  stat_compare_means(comparisons = compar.stbs)+
  #stat_summary() +
  geom_jitter(alpha=0.5, size=3, width = 0.2) +
  theme_minimal()
core_domdf <- meta(ps.main.rar)
ggplot(core_domdf, aes(condition_status,core_dom)) +
  geom_boxplot() + 
  stat_compare_means(comparisons = compar.stbs)+
  #stat_summary() +
  geom_jitter(alpha=0.5, size=3, width = 0.2) +
  theme_minimal()


```



# Variability   

```{r}

ps.merge <- readRDS("data/rds/ps.merge.rds")
ps.merge.genus.rel <- microbiome::transform(tax_glom(ps.merge,"Genus"), "compositional")
taxa_names(ps.merge.genus.rel) <- tax_table(ps.merge.genus.rel)[,"Genus"]
table(tax_table(ps.merge.genus.rel)[,"Phylum"])
```



```{r fig.height=6, fig.width=12}
#ps.main.gen.rel <- readRDS("ms/rds/ps.main.gen.rel.rds")
prevalences <- seq(.05, 1, .05)
detections <- 10^seq(log10(1e-5), log10(.10), length = 10)

library(RColorBrewer)
core_gen_p <- plot_core(ps.merge.genus.rel,  prevalences=prevalences,
                        detections=detections, 
                        min.prevalence = 0.75, 
                        plot.type = "heatmap",
                        colours = rev(brewer.pal(5, "Spectral")),
                        horizontal = T) + 
  theme_bw() +
  theme(axis.text.x = element_text(face="italic")) + 
  rotate_x_text() +
  labs(subtitle = paste0("Core nasopharyngeal genera (n = ", nsamples(ps.merge.genus.rel), ")"))
core_gen_p

0.00001

detections2 <- c(0.00001, 0.0001, 0.001, 0.01, 0.1)
core_gen_p_vertical <- plot_core(ps.merge.genus.rel,  prevalences=prevalences,
                        detections=detections2, 
                        min.prevalence = 0.75, 
                        plot.type = "heatmap",
                        colours = rev(brewer.pal(5, "Spectral")),
                        horizontal = F) + 
  theme_bw() +
  theme(axis.text.y = element_text(face="italic")) + 
  rotate_x_text() +
  labs(subtitle = paste0("Core nasopharyngeal genera)"))
core_gen_p_vertical

```


```{r fig.width=6, fig.height=4}
plot_core(ps.merge.genus.rel)
```

## Plot Core Taxa Tree  

```{r}

core_gen <- core_members(ps.merge.genus.rel, detection = 0.00001, prevalence = 0.75)

ps_core_gen <- core(ps.merge.genus.rel, detection = 0.00001, prevalence = 0.75)

```



```{r}
#taxa_names(ps.merge.genus.rel) <- tax_table(ps.merge.genus.rel)[,"Genus"]
core_gen <- core_members(ps.merge.genus.rel, detection = 0.00001, prevalence = 0.75)

ps_core_gen <- core(ps.merge.genus.rel, detection = 0.00001, prevalence = 0.75)

no_ili_ps <- subset_samples(ps_core_gen, condition_status=="control")
no_ili_ps <- prune_taxa(taxa_sums(no_ili_ps) > 0 , no_ili_ps)

no_ili_14d_ps <- subset_samples(ps_core_gen, condition_status=="control_14_days")
no_ili_14d_ps <- prune_taxa(taxa_sums(no_ili_14d_ps) > 0 , no_ili_14d_ps)

acut_ili_ps <- subset_samples(ps_core_gen, condition_status=="acute_ili")

#rand_ili_samples <- sample(sample_names(acut_ili_ps), 98)
#acut_ili_ps <- prune_samples(sample_names(acut_ili_ps) %in% rand_ili_samples, acut_ili_ps)
acut_ili_ps <- prune_taxa(taxa_sums(acut_ili_ps) > 0 , acut_ili_ps)

ili_14d_ps <- subset_samples(ps_core_gen, condition_status=="acute_ili_14_days")
ili_14d_ps <- prune_taxa(taxa_sums(ili_14d_ps) > 0 , ili_14d_ps)


recov_ili_ps <- subset_samples(ps_core_gen, condition_status=="acute_ili_recovery")
recov_ili_ps <- prune_taxa(taxa_sums(recov_ili_ps) > 0 , recov_ili_ps)

#no_ili','no_ili_14_days','acute_ili',                                        'acute_ili_14_days','acute_ili_recovery

```


### check for taxa variability  
#### Calculate PV  

```{r}
set.seed(1215)
no_illi_pv <- taxa_pv_boot(no_ili_ps,bs.iter=999)  
no_ili_pv2 <- no_illi_pv %>% 
  dplyr::select(-starts_with("PVI")) %>% 
  mutate(Condition = "control") %>% 
  arrange(desc(meanPV))


no_ili_14d_pv <- taxa_pv_boot(no_ili_14d_ps,bs.iter=999)  
no_ili_14d_pv2 <- no_ili_14d_pv %>% 
  dplyr::select(-starts_with("PVI")) %>% 
  mutate(Condition = "control_14_days") %>% 
  arrange(desc(meanPV))


acut_ili_pv <- taxa_pv_boot(acut_ili_ps,bs.iter=999)  
acut_ili_pv2 <- acut_ili_pv %>% 
  dplyr::select(-starts_with("PVI")) %>% 
  mutate(Condition = "acute_ili") %>% 
  arrange(desc(meanPV))


ili_14d_pv <- taxa_pv_boot(ili_14d_ps,bs.iter=999)  
ili_14d_pv2 <- ili_14d_pv %>% 
  dplyr::select(-starts_with("PVI")) %>% 
  mutate(Condition = "acute_ili_14_days") %>% 
  arrange(desc(meanPV))


recov_ili_pv <- taxa_pv_boot(recov_ili_ps, bs.iter=999)  
recov_ili_pv2 <- recov_ili_pv %>% 
  dplyr::select(-starts_with("PVI")) %>% 
  mutate(Condition = "acute_ili_recovery") %>% 
  arrange(desc(meanPV))
#recov_ili_pv2
#no_illi_pv
```

#### Plot core taxa PV  

```{r fig.width=14, fig.height=10}
core_pv <- bind_rows(no_ili_pv2,
                     no_ili_14d_pv2,
                     acut_ili_pv2,
                     ili_14d_pv2,
                     recov_ili_pv2)

core_pv$LowerCI <- as.numeric(core_pv$LowerCI)
core_pv$UpperCI <- as.numeric(core_pv$UpperCI)

saveRDS(core_pv, "data/rds/core_pv.rds")

pv_plot <- ggplot(core_pv, aes(x = reorder(Taxa,-meanPV), y = meanPV)) +
  #geom_point(aes(color = Condition)) +
  geom_pointrange(aes(ymax=UpperCI, ymin=LowerCI, 
                      color = Condition), 
                  alpha=0.50)  + 
  scale_color_manual(values = dis_col) + 
  coord_flip() + 
  theme_biome_utils() +
  ylim(0.4,0.9) +
  theme(axis.text.y = element_text(face="italic")) +
  xlab("Genus") +
  ylab("Proportional variability (mean)")
pv_plot

core_gen_p 

core_gen_p + pv_plot + plot_layout(h=c(1,2)) + plot_annotation(tag_levels = "A")
ggsave("05_stability/coreplot_taxa_pv.pdf", height = 6, width = 10)
ggsave("05_stability/coreplot_taxa_pv.tiff", height = 6, width = 10)

write.table(core_pv, "05_stability/pv_core_taxa.txt", sep="\t")
```



```{r}
stab.plot.f <-  stab.plot + theme_minimal(base_size = 10) + theme(legend.position="none")
stab.line.plot.f <- stab.line.plot + theme_minimal(base_size = 10)
p.dv.f <- p.dv + theme_minimal(base_size = 10)

(stab.plot.f+stab.line.plot.f)/p.dv.f + plot_annotation(tag_levels = "A")
ggsave("msfigs/stability.bet.alpha.spearman.cor.pdf", height = 6, width = 10)

```

## Distirbution of PV 

```{r}

no_illi_pv3  <- no_illi_pv  %>% 
  dplyr::select(starts_with("PVI")) %>% 
  rownames_to_column("taxa") %>% 
  reshape2::melt() %>% 
  mutate(Condition = "control")

no_ili_14d_pv3  <- no_ili_14d_pv  %>% 
  dplyr::select(starts_with("PVI")) %>% 
  rownames_to_column("taxa") %>% 
  reshape2::melt() %>% 
  mutate(Condition = "control_14_days")

acut_ili_pv3 <- ili_14d_pv  %>% 
  dplyr::select(starts_with("PVI")) %>% 
  rownames_to_column("taxa") %>% 
  reshape2::melt() %>% 
  mutate(Condition = "acute_ili")

ili_14d_pv3 <- ili_14d_pv  %>% 
  dplyr::select(starts_with("PVI")) %>% 
  rownames_to_column("taxa") %>% 
  reshape2::melt() %>% 
  mutate(Condition = "acute_ili_14_days")


recov_ili_pv3 <- recov_ili_pv  %>% 
  dplyr::select(starts_with("PVI")) %>% 
  rownames_to_column("taxa") %>% 
  reshape2::melt() %>% 
  mutate(Condition = "acute_ili_recovery")

core_pv3 <- bind_rows(no_illi_pv3,
                      no_ili_14d_pv3,
                      acut_ili_pv3,
                      ili_14d_pv3,
                      recov_ili_pv3)

#head(recov_ili_pv3)

ridge_pv <- ggplot(core_pv3, aes(x = value, y = Condition)) + 
  geom_density_ridges(aes(fill=Condition), alpha=0.75, color=NA)+
  scale_fill_manual(values = dis_col) +
  theme_minimal() +
  facet_wrap(~taxa, ncol=3) +
  theme(axis.text.y = element_blank(),
        axis.title.y= element_blank(),
        strip.text = element_text(face="italic"))+
  xlab("Proportional variability")


```

```{r}

core_gen_p_vertical <- core_gen_p_vertical + theme_minimal(base_size = 10)

core_gen_p_vertical + ridge_pv + plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect", widths = c(1,2))

ggsave("msfigs/Figure core pv_density.pdf", height = 5, width = 10)
```


## Extract ASVs within core genera  

```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(ggtree)

ps.merge.clr <- microbiome::transform(ps.merge,"clr")
ps.core.asv <- subset_taxa(ps.merge.clr, Genus %in% core_gen)
ps.core.asv <- subset_samples(ps.core.asv, condition_status %in% c("control", "acute_ili"))
ps.core.asv <- removeZeroTaxa(ps.core.asv)

colnames(sample_data(ps.core.asv))[2] <- "samid"
taxa_fd <- calculateTaxaFoldDifference(ps.core.asv, group="condition_status") 

treedata<-taxa_fd %>%
  filter(abs(FoldDifference) > 0.3 & abs(FoldDifference) !=Inf) 

tree <- ps.core.asv@phy_tree
tree <- drop.tip(tree, tree$tip.label[!tree$tip.label %in% treedata$FeatureID])

#ggtree(tree, layout="circular",
#       aes(color=Phylum)) %<+%
#  treedata
library(RColorBrewer)
testcol <-  colorRampPalette( brewer.pal(12, "Paired"))
tax_cols <- c("#e6194b","brown3","#7789c9","#3cb44b","#f58231","#911eb4","#0dbfbf","#b097ad","#bcf60c","#fabebe","#008080","#e6beff","#9a6324","#e4cd05","#800000","#88cf9e","#808000")
#,                     size=Prevalence.acute_ili


## Select those with high low
treedata2 <- treedata %>% 
  filter(abs(FoldDifference) > 0.3 | abs(FoldDifference) > 0.3)
  
d1 <- data.frame(x = seq(1.8, 3.5, length.out = 9))
p <- ggtree(tree, layout="circular") %<+%
  treedata +
  #geom_tree(aes(color=Genus), size=2) +
  geom_point2(aes(color=Genus), size=3, shape=21) +
  #geom_tippoint(aes(fill=Enriched), 
  #             shape=21, 
  #alpha=0.5,
  #            size=2) +
  #geom_tiplab(aes(label=Significant)) 
  geom_tiplab(aes(color=Enriched),offset = .30, align=TRUE, linesize=0.5, size=3) +
  scale_color_manual(values = tax_cols) +
  #scale_fill_manual(values = dis_col) +
  geom_tippoint(aes(fill=FoldDifference), x = d1$x[1], shape = 21,
                na.rm=TRUE,color="#999999", size=3) +
  scale_fill_gradient2() +
  theme(legend.text = element_text(face="italic"))
p

ggsave("msfigs/Figure coretee ctrl iliv1.pdf", height = 6, width = 9)

```


```{r}
sessionInfo()
```


