---
title: "The nasopharyngeal microbiota of elderly subjects with and without influenza-like illness"
subtitle: "Taxa association"
author: "Sudarshan A. Shetty"
date: "`r date()`"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
    toc_float: true
---

# Association analysis  

## Libraries

```{r message=FALSE, warning=FALSE}
library(microbiome)
library(microbiomeutilities)
library(phyloseq)
library(coin)
library(dplyr)
library(readr)

dis_col <- c(acute_ili="#6883ba", control = "#166f65", `NA`="black",
             acute_ili_14_days = "#6d597a", control_14_days = "#87d2c9", 
             acute_ili_recovery = "#e56b6f") 
```

## Import functions  
```{r}
#source("codes/functions.associations.R")
#source("codes/transform_to_fractions.R")
source("codes/biomestats.R") # all function for this specific analysis are stored here.
source("codes/generalCodes.R")

```

```{r}
dir.create("06_taxa_associations")
```


## Import data  
```{r eval=TRUE}

ps.main.2 <- readRDS("data/rds/ps.main.2.rds")

colnames(tax_table(ps.main.2))[7] <- "Species"
ps.main.2.genus <- tax_glom(ps.main.2,"Genus", NArm = FALSE)
saveRDS(ps.main.2.genus, "data/rds/ps.main.2.genus.rds")

#DT::datatable(tax_table(ps.main.2.genus))
#ps.main.2 <- format_to_besthit(ps.main.2)
#taxa_names(ps.main.2)[3:15]
#ps.main.genus.2 <- readRDS("data/rds/ps.main.genus.2.rds")
```

```{r}
ps.main.2.genus <- readRDS("data/rds/ps.main.2.genus.rds")
#ps.main.2.genus <- readRDS("data/rds/ps.main.2.genus.rds")
#table(tax_table(ps.main.2.genus)[,"Genus"], useNA="always")

ps.main.2.genus <- format_to_besthit(ps.main.2.genus)

taxa_names(ps.main.2.genus)[3:15]

taxa_names(ps.main.2.genus) <- gsub(" ", ".", taxa_names(ps.main.2.genus))
taxa_names(ps.main.2.genus) <- gsub("\\[|\\]", "", taxa_names(ps.main.2.genus))
taxa_names(ps.main.2.genus) <- gsub("-", ".", taxa_names(ps.main.2.genus))
```


### Variables of interest:  
Antibiotics in 2014, Antimicrobials in 2014, Co-morbidities 2014 

```{r} 
main_var <- "ILI"

comord.vars <- c("BMI_2014","respiratory_disease_2014", "comorbidities_2014",
                 "chronic_cvd_2014","Diabetes_2014", "autoimmune_disease_2014", 
                 "other_malignancies", "Roken_2014")

med.vars <- c("ace_inhibitors_2014", "statins_2014",
              "alpa_beta_blockers_2014","antimicrobials_2014", "arbs_2014", "insulin_2014", "ppi_2014", 
              "Corticosteroiden_2014", "Antiepileptica_2014", "Antihistaminica_2014",
              "analgesics_2014", "flu_shot_2014_2015", "denzodiazepine_2014",
              "Antidepressiva_2014", "Calciumregulatie_2014",
              "AstmaCOPD_2014", "number_of_medications_2014", "antibioitics_2014")
# removed "Dopaminergica_2014" due to very low numbers
# This will be used only when comparing ILI participants
#patho.vars <- c("influenza_2014_any","haemophilus_2014_any","coronavirus_2014",
#                "respiratory_syncytial_virus_2014","staphylococcus_aureus_2014",
#                "haemolytic_streptococcus_2014","human_metapneumovirus_2014",
#                "human_parainfluenzavirus_2014_any","rhinovirus_2014")

#all.vars <- c(comord.vars,med.vars,patho.vars)
all.vars <- c(comord.vars,med.vars)

```

```{r}
#all.vars
#sample_data(ps.main.genus.2) <- sample_data(ps.main.genus.2)[,all.vars]

table(sample_data(ps.main.2.genus)$ppi_2014, useNA = "always")

```

1= participant is smoker
2= participant is passive smoker
3= participant doesn't smoke
```{r}


table(sample_data(ps.main.2.genus)$Roken_2014, useNA = "always")
# Roken i.e. smoking has three levels. 1=smoking, 2=passive smoking and 3-not smoking

# Recode smoking variables: so 0 is no and increasing to yes 2
sample_data(ps.main.2.genus)$Roken_2014 <- 3-as.numeric(sample_data(ps.main.2.genus)$Roken_2014)

table(sample_data(ps.main.2.genus)$flu_shot_2014_2015, useNA = "always")
sample_data(ps.main.2.genus)[,"flu_shot_2014_2015"][sample_data(ps.main.2.genus)[,"flu_shot_2014_2015"] == 3] <- 2

table(sample_data(ps.main.2.genus)$antibioitics_2014, useNA = "always") 
```


# Control-ILI    

## Prep abundance data  
```{r}
dir.create("06_taxa_associations/ctrl_ili")

ctrl_ili_v1 <- c("control", "acute_ili")

ps.ctrl.iliv1 <- subset_samples(ps.main.2.genus, 
                                condition_status %in% ctrl_ili_v1)

# clean taxonomy
ps.ctrl.iliv1 <- prune_taxa(taxa_sums(ps.ctrl.iliv1) > 0, ps.ctrl.iliv1)
```


## Create taxa abundance table 
```{r}

asv.ctrl.iliv1 <- transform_to_fractions(ps.ctrl.iliv1, 
                                         det.thres = 0.001,
                                         prev.thres = 10/100)
dim(asv.ctrl.iliv1)
#colnames(asv.ctrl.iliv1)
#asv.tab.ctrl.ili[1:5,1:5]
```


## Prep metadata  

```{r}

pheno.data.v1 <- prep_metadata(ps.ctrl.iliv1, 
                               sam.ids = "Sample_Name", 
                               ref.group="control", 
                               compare= "condition_status", 
                               confounder.vars = c("gender", "age_yrs_oct2014"),
                               var.interest=all.vars)

#head(pheno.data.v1)
```


## Strata frequency  
```{r eval=TRUE}
frequency.of.strata <- table(pheno.data.v1$stratum); mean(frequency.of.strata); frequency.of.strata
```

## Data for testing
```{r}
# remove firt column from asv_dat it is subject id
asv.ctrl.iliv1.dat <- dplyr::select(asv.ctrl.iliv1, -ID) 
# Create "data.set":
data.set.v1 <- cbind(pheno.data.v1,asv.ctrl.iliv1.dat)
row.names(data.set.v1) <- NULL
# get only bacterial names
ncols.pheno.v1 <- ncol(pheno.data.v1) +1

nocl.v1 <- ncol(data.set.v1)-5
bacterial.abund.v1 <- names(data.set.v1)[ ncols.pheno.v1:nocl.v1]

other <- c("average","SD","prop.positive","entropy","CV")

#vars.test <- colnames(phenotypic.data.1)[c(2,5:39)]
vars.test.v1 <- c("ILI",all.vars)

data.for.testing.v1 <- subset(data.set.v1,
                              select=c(vars.test.v1,bacterial.abund.v1,other,"stratum"))
#head(data.for.testing)
#saveRDS(data.for.testing, "ms/taxa_test/ctrl_ili/data.for.testing.ctrl.ili.rds")
```

## Type of vars 
```{r eval=TRUE}

TYPES.v1 <- c("ordinal","continuous",rep("ordinal",nocl.v1),
           rep("continuous",5),
           "categorical")

types.v1 <- c("categorical","continuous", 
           rep("categorical",length(vars.test.v1)-2),
           rep("dirac.and.continuous",length(bacterial.abund.v1)),
           rep("continuous",5),
           "categorical")
#colnames(data.for.testing)[1:10]

```


## Test associations   
```{r}

no.of.factors <- length(vars.test.v1)
aux.columns <- (length(vars.test.v1)+1):(length(colnames(data.for.testing.v1))-6)
variables.of.interest <- names(data.for.testing.v1)
#nominal.bound.on.FDR <- 0.20
log.scale <- TRUE

```


```{r eval=TRUE}
#dir.create("../../test_stat")
set.seed(1210125)
check_association(data.for.testing.v1,
                  B = 100000,
                  no.of.factors= no.of.factors,
                  aux.columns = aux.columns,
                  log.scale = TRUE,
                  nominal.bound.on.FDR = 0.20,
                  phen.data = pheno.data.v1,
                  compare.label="ctrl_ili",
                  aux.TYPES=TYPES.v1,
                  cat.types = types.v1,
                  path_loc="06_taxa_associations/ctrl_ili/",
                  name.of.stratification= "Sex and age-group")
```


## Visualize assocaitions  
```{r eval=TRUE, message=FALSE}

path.files <- "06_taxa_associations/ctrl_ili/"
pattern.name <- "SELECTED.associations.between.ctrl_ili."
test.ctrl.ili <- combine_tables(path.files,pattern.name)

```

```{r}
test.ctrl.ili$variable <- gsub("_2014","", test.ctrl.ili$variable)

p <- plot_taxa_associations(tab=test.ctrl.ili, 
                       ps=ps.ctrl.iliv1, 
                       adj.pval=TRUE, 
                       group="condition_status",
                       cut.off = 0.05,
                       p.widths = c(1,0.5)) 
p <- p + theme(text = element_text(size=10))

ggsave("06_taxa_associations/ctrl_ili/ctrl_ili_taxa_associations.pdf", h=6,w=8)
```


```{r eval=T, message=FALSE}

save_xlsx_workbook(path.files="06_taxa_associations/ctrl_ili/",
                   pattern.name="SELECTED.associations.between.ctrl_ili.")

```

### Plot significant taxa  
```{r message=FALSE, warning=FALSE,eval=TRUE}
path.files <- "06_taxa_associations/ctrl_ili/"
pattern.name <- "SELECTED.associations.between.ctrl_ili."
all_test <- combine_tables(path.files,pattern.name)
test_ili <- subset(all_test, variable == "ILI" & better.bound.FDR <= 0.05)

p.ili <- plot_listed_taxa(ps.ctrl.iliv1, select.taxa = test_ili$taxa,
                          group.order = c("control","acute_ili"),
                          group = "condition_status",
                          group.colors = dis_col) + 
  ylab("Relative abundance (log10)") + 
  scale_y_log10() +
  theme(strip.text = element_text(face = "italic"),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())
  
p.ili

ggsave("06_taxa_associations/ctrl_ili/ctrl_ili_taxa.pdf", h=3,w=6)
```

```{r}
ctrl.test.ctrl.ili <- test.ctrl.ili
ctrl.test.ctrl.ili$Association <- ifelse(ctrl.test.ctrl.ili$sign == -1, "Negative", "Positive")
ctrl.test.ctrl.ili <- ctrl.test.ctrl.ili %>% 
    filter(p.value <= 0.05)

DT::datatable(ctrl.test.ctrl.ili)
```



```{r}
p / p.ili
```


# Control vs ILI v2 
 acute_ili  acute_ili_14_days acute_ili_recovery            control 
               150                188                180                160 
```{r}
#dir.create("06_taxa_associations/ctrl_ili")

ctrl_ili_v2 <- c("control", "acute_ili_14_days")

ps.ctrl.iliv2 <- subset_samples(ps.main.2.genus, 
                                condition_status %in% ctrl_ili_v2)

# clean taxonomy
ps.ctrl.iliv2 <- prune_taxa(taxa_sums(ps.ctrl.iliv2) > 0, ps.ctrl.iliv2)


```

## Create taxa abundance table 
```{r}

#length(core_members(transform(ps.genus.ctrl.ili, "compositional"), 0.001, 5/100))


asv.ctrl.iliv2 <- transform_to_fractions(ps.ctrl.iliv2, 
                                           det.thres = 0.001,
                                           prev.thres = 10/100)
dim(asv.ctrl.iliv2)
#asv.tab.ctrl.ili[1:5,1:5]
## Prep metadata  

pheno.1v2 <- prep_metadata(ps.ctrl.iliv2, 
                                     sam.ids = "Sample_Name", 
                                     ref.group="control", 
                                     compare= "condition_status", 
                                     confounder.vars = c("gender", "age_yrs_oct2014"),
                                     var.interest=all.vars)

#head(pheno.1v2)


## Data for testing

# remove firt column from asv_dat it is subject id
asv.ctrl.iliv2.dat <- dplyr::select(asv.ctrl.iliv2, -ID) 
# Create "data.set":
data.set.1.v2 <- cbind(pheno.1v2,asv.ctrl.iliv2.dat)
row.names(data.set.1.v2) <- NULL
# get only bacterial names
ncols.pheno.v2 <- ncol(pheno.1v2) +1
nocl.v2 <- ncol(data.set.1.v2)-5
bacterial.abund.v2 <- names(data.set.1.v2)[ ncols.pheno.v2:nocl.v2]

other <- c("average","SD","prop.positive","entropy","CV")

#vars.test <- colnames(phenotypic.data.1)[c(2,5:39)]
vars.test.v2 <- c("ILI")

data.for.testing.v2 <- subset(data.set.1.v2,
                           select=c(vars.test.v2,bacterial.abund.v2,other,"stratum"))
#head(data.for.testing)
#saveRDS(data.for.testing, "ms/taxa_test/ctrl_ili/data.for.testing.ctrl.ili.rds")
```

## Type of vars 
```{r eval=TRUE}

TYPES.v2 <- c("ordinal",rep("ordinal",nocl.v2),
           rep("continuous",5),
           "categorical")

types.v2 <- c("categorical", 
           rep("dirac.and.continuous",length(bacterial.abund.v2)),
           rep("continuous",5),
           "categorical")
#colnames(data.for.testing)[1:10]

```


## Test associations   
```{r}
no.of.factors <- length(vars.test.v2)
aux.columns <- (length(vars.test.v2)+1):(length(colnames(data.for.testing.v2))-6)
variables.of.interest <- names(data.for.testing.v2)
#nominal.bound.on.FDR <- 0.20
log.scale <- TRUE

```


```{r eval=TRUE}
dir.create("06_taxa_associations/ctrl_iliv2/")
set.seed(12136125)
check_association(data.for.testing.v2,
                  B = 100000,
                  no.of.factors= no.of.factors,
                  aux.columns = aux.columns,
                  log.scale = TRUE,
                  nominal.bound.on.FDR = 0.20,
                  phen.data = pheno.1v2,
                  compare.label="ctrl_iliv2",
                  aux.TYPES=TYPES.v2,
                  cat.types = types.v2,
                  path_loc="06_taxa_associations/ctrl_iliv2/",
                  name.of.stratification= "Sex and age-group")
```


```{r}
path.files <- "06_taxa_associations/ctrl_iliv3/"
pattern.name <- "SELECTED.associations.between."
ctrl_iliv2.iliv1 <- combine_tables(path.files,pattern.name)
ctrl_iliv2.iliv1$Association <- ifelse(ctrl_iliv2.iliv1$sign == -1, "Negative", "Positive")
ctrl_iliv2.iliv1 <- ctrl_iliv2.iliv1 %>% 
    filter(p.value <= 0.05)

DT::datatable(ctrl_iliv2.iliv1)
```

## Control vs ILI v3  

 
```{r}
#dir.create("06_taxa_associations/ctrl_ili")

ctrl_ili_v3 <- c("control", "acute_ili_recovery")

ps.ctrl.iliv3 <- subset_samples(ps.main.2.genus, 
                                condition_status %in% ctrl_ili_v3)

# clean taxonomy
ps.ctrl.iliv3 <- prune_taxa(taxa_sums(ps.ctrl.iliv3) > 0, ps.ctrl.iliv3)


```

## Create taxa abundance table 
```{r}

#length(core_members(transform(ps.genus.ctrl.ili, "compositional"), 0.001, 5/100))


asv.ctrl.iliv3 <- transform_to_fractions(ps.ctrl.iliv3, 
                                           det.thres = 0.001,
                                           prev.thres = 10/100)
dim(asv.ctrl.iliv3)
#asv.tab.ctrl.ili[1:5,1:5]
## Prep metadata  

pheno.1v3 <- prep_metadata(ps.ctrl.iliv3, 
                           sam.ids = "Sample_Name", 
                           ref.group="control", 
                           compare= "condition_status", 
                           confounder.vars = c("gender", "age_yrs_oct2014"),
                           var.interest=all.vars)

#head(pheno.1v3)


## Data for testing

# remove firt column from asv_dat it is subject id
asv.ctrl.iliv3.dat <- dplyr::select(asv.ctrl.iliv3, -ID) 
# Create "data.set":
data.set.1.v3 <- cbind(pheno.1v3,asv.ctrl.iliv3.dat)
row.names(data.set.1.v3) <- NULL
# get only bacterial names
ncols.pheno.v3 <- ncol(pheno.1v3) +1
nocl.v3 <- ncol(data.set.1.v3)-5
bacterial.abund.v3 <- names(data.set.1.v3)[ ncols.pheno.v3:nocl.v3]

other <- c("average","SD","prop.positive","entropy","CV")

#vars.test <- colnames(phenotypic.data.1)[c(2,5:39)]
vars.test.v3 <- c("ILI")

data.for.testing.v3 <- subset(data.set.1.v3,
                           select=c(vars.test.v3,bacterial.abund.v3,other,"stratum"))
#head(data.for.testing)
#saveRDS(data.for.testing, "ms/taxa_test/ctrl_ili/data.for.testing.ctrl.ili.rds")
```

## Type of vars 
```{r eval=TRUE}

TYPES.v3 <- c("ordinal",rep("ordinal",nocl.v3),
           rep("continuous",5),
           "categorical")

types.v3 <- c("categorical", 
           rep("dirac.and.continuous",length(bacterial.abund.v3)),
           rep("continuous",5),
           "categorical")
#colnames(data.for.testing)[1:10]

```


## Test associations   
```{r}
no.of.factors <- length(vars.test.v3)
aux.columns <- (length(vars.test.v3)+1):(length(colnames(data.for.testing.v3))-6)
variables.of.interest <- names(data.for.testing.v3)
#nominal.bound.on.FDR <- 0.20
log.scale <- TRUE

```


```{r eval=TRUE}
dir.create("06_taxa_associations/ctrl_iliv3/")
set.seed(12136125)
check_association(data.for.testing.v3,
                  B = 100000,
                  no.of.factors= no.of.factors,
                  aux.columns = aux.columns,
                  log.scale = TRUE,
                  nominal.bound.on.FDR = 0.20,
                  phen.data = pheno.1v3,
                  compare.label="ctrl_iliv3",
                  aux.TYPES=TYPES.v3,
                  cat.types = types.v3,
                  path_loc="06_taxa_associations/ctrl_iliv3/",
                  name.of.stratification= "Sex and age-group")
```


```{r}
path.files <- "06_taxa_associations/ctrl_iliv3/"
pattern.name <- "SELECTED.associations.between."
ctrl_iliv3.iliv1 <- combine_tables(path.files,pattern.name)
ctrl_iliv3.iliv1$Association <- ifelse(ctrl_iliv3.iliv1$sign == -1, "Negative", "Positive")
ctrl_iliv3.iliv1 <- ctrl_iliv3.iliv1 %>% 
    filter(p.value <= 0.05)

DT::datatable(ctrl_iliv3.iliv1)
```

# Check for ILI associations with pathogens

```{r}


patho.vars <- c("influenza_2014_any","haemophilus_2014_any","coronavirus_2014",
                "respiratory_syncytial_virus_2014","staphylococcus_aureus_2014",
                "haemolytic_streptococcus_2014","human_metapneumovirus_2014",
                "human_parainfluenzavirus_2014_any","rhinovirus_2014")
all.vars <- c(comord.vars,med.vars,patho.vars)
```

## ILI V1 vs pathogens and other data  
```{r}

ps.iliv1 <- subset_samples(ps.main.2.genus, 
                         condition_status == "acute_ili")

# clean taxonomy
ps.iliv1 <- prune_taxa(taxa_sums(ps.iliv1) > 0, ps.iliv1)

```


## Create taxa abundance table 
```{r}

asv.iliv1 <- transform_to_fractions(ps.iliv1, 
                                    det.thres = 0.001,
                                    prev.thres = 10/100)
dim(asv.iliv1)
#asv.tab.ctrl.ili[1:5,1:5]
```


## Prep metadata  

```{r}

pheno.data.ili.v1 <- prep_metadata(ps.iliv1, 
                                   sam.ids = "Sample_Name", 
                                   ref.group="acute_ili", 
                                   compare= "condition_status", 
                                   confounder.vars = c("gender", "age_yrs_oct2014"),
                                   var.interest=all.vars)

#head(pheno.data.ili.v1)

pheno.data.ili.v1$human_metapneumovirus_2014[pheno.data.ili.v1$human_metapneumovirus_2014==3] <- 2
pheno.data.ili.v1$respiratory_syncytial_virus_2014[pheno.data.ili.v1$respiratory_syncytial_virus_2014==3] <- 2

pheno.data.ili.v1$human_parainfluenzavirus_2014_any[pheno.data.ili.v1$human_parainfluenzavirus_2014_any==3] <- 2

pheno.data.ili.v1$rhinovirus_2014[pheno.data.ili.v1$rhinovirus_2014==3] <- 2
```


## Strata frequency  
```{r eval=TRUE}
frequency.of.strata <- table(pheno.data.ili.v1$stratum); mean(frequency.of.strata); frequency.of.strata
```

## Data for testing
```{r}
# remove firt column from asv_dat it is subject id
asv.iliv1.dat <- dplyr::select(asv.iliv1, -ID) 
# Create "data.set":
data.set.ili.v1 <- cbind(pheno.data.ili.v1,asv.iliv1.dat)

row.names(data.set.ili.v1) <- NULL
# get only bacterial names

#ncols.pheno.v1 <- ncol(pheno.data.v1) +1
#nocl.v1 <- ncol(data.set.v1)-5

ncols.pheno.ili.v1 <- ncol(pheno.data.ili.v1) +1

nocl.ili.v1 <- ncol(data.set.ili.v1)-5
bacterial.abund.ili.v1 <- names(data.set.ili.v1)[ ncols.pheno.ili.v1:nocl.ili.v1]

other <- c("average","SD","prop.positive","entropy","CV")

#vars.test <- colnames(phenotypic.data.1)[c(2,5:39)]
vars.test.ili.v1 <- c(all.vars)

data.for.testing.ili.v1 <- subset(data.set.ili.v1,
                              select=c(vars.test.ili.v1,
                                       bacterial.abund.ili.v1,
                                       other,"stratum"))
#head(data.for.testing)
#saveRDS(data.for.testing, "ms/taxa_test/ctrl_ili/data.for.testing.ctrl.ili.rds")
```

## Type of vars 
```{r eval=TRUE}

TYPES.ili.v1 <- c("continuous",rep("ordinal",nocl.ili.v1),
           rep("continuous",5),
           "categorical")

types.ili.v1 <- c("continuous", 
           rep("categorical",length(vars.test.ili.v1)-2),
           rep("dirac.and.continuous",length(bacterial.abund.ili.v1)),
           rep("continuous",5),
           "categorical")
#colnames(data.for.testing)[1:10]

```


## Test associations   
```{r}

no.of.factors <- length(vars.test.ili.v1)
aux.columns <- (length(vars.test.ili.v1)+1):(length(colnames(data.for.testing.ili.v1))-6)
variables.of.interest <- names(data.for.testing.ili.v1)
#nominal.bound.on.FDR <- 0.20
log.scale <- TRUE

```


```{r eval=TRUE}
#dir.create("../../test_stat")
dir.create("06_taxa_associations/iliv1/")
set.seed(1210125)
check_association(data.for.testing.ili.v1,
                  B = 100000,
                  no.of.factors= no.of.factors,
                  aux.columns = aux.columns,
                  log.scale = TRUE,
                  nominal.bound.on.FDR = 0.20,
                  phen.data = pheno.data.ili.v1,
                  compare.label="ili_v1",
                  aux.TYPES=TYPES.ili.v1,
                  cat.types = types.ili.v1,
                  path_loc="06_taxa_associations/iliv1/",
                  name.of.stratification= "Sex and age-group")
```


```{r}
path.files <- "06_taxa_associations/iliv1/"
pattern.name <- "SELECTED.associations.between."
test.iliv1 <- combine_tables(path.files,pattern.name)
test.iliv1$Association <- ifelse(test.iliv1$sign == -1, "Negative", "Positive")
test.iliv1 <- test.iliv1 %>% 
    filter(p.value <= 0.05)

DT::datatable(test.iliv1)
```

# Check ILI V2 vs Pathogens  

```{r}

ps.iliv2 <- subset_samples(ps.main.2.genus, 
                         condition_status == "acute_ili_14_days")

# clean taxonomy
ps.iliv2 <- prune_taxa(taxa_sums(ps.iliv2) > 0, ps.iliv2)

```


## Chnage metadata

Add infection information from ili v1 to ili v2 microbiota 
```{r}

ili.v1.meta <- get_tibble(ps.iliv1, slot="sam_data")

ps.iliv2.sub <- subset_samples(ps.iliv2, participant_id %in% ili.v1.meta$participant_id)

ili.v1.meta <- ili.v1.meta %>% 
  filter(participant_id %in% sample_data(ps.iliv2.sub)$participant_id) %>% 
  as.data.frame() %>% 
  column_to_rownames("participant_id")

sample_names(ps.iliv2.sub) <- rownames(ili.v1.meta)
sample_data(ps.iliv2.sub) <- sample_data(ili.v1.meta)

table(sample_data(ps.iliv2.sub)$condition_status, sample_data(ps.iliv2.sub)$influenza_2014_any)
table(sample_data(ps.iliv2)$condition_status, sample_data(ps.iliv2)$haemophilus_2014_any)

ps.iliv2 <- ps.iliv2.sub
```


## Create taxa abundance table 
```{r}

asv.iliv2 <- transform_to_fractions(ps.iliv2, 
                                    det.thres = 0.001,
                                    prev.thres = 10/100)
dim(asv.iliv2)
#asv.tab.ctrl.ili[1:5,1:5]
```


## Prep metadata  

```{r}
#human_metapneumovirus_2014
#respiratory_syncytial_virus_2014
#human_parainfluenzavirus_2014_any 
#rhinovirus_2014


pheno.data.ili.v2 <- prep_metadata(ps.iliv2, 
                                   sam.ids = "Sample_Name", 
                                   ref.group="acute_ili", 
                                   compare= "condition_status", 
                                   confounder.vars = c("gender", "age_yrs_oct2014"),
                                   var.interest=all.vars)

head(pheno.data.ili.v2)


pheno.data.ili.v2$human_metapneumovirus_2014[pheno.data.ili.v2$human_metapneumovirus_2014==3] <- 2

pheno.data.ili.v2$human_metapneumovirus_2014[pheno.data.ili.v2$human_metapneumovirus_2014==3] <- 2
pheno.data.ili.v2$respiratory_syncytial_virus_2014[pheno.data.ili.v2$respiratory_syncytial_virus_2014==3] <- 2

pheno.data.ili.v2$human_parainfluenzavirus_2014_any[pheno.data.ili.v2$human_parainfluenzavirus_2014_any==3] <- 2

pheno.data.ili.v2$rhinovirus_2014[pheno.data.ili.v2$rhinovirus_2014==3] <- 2

#summary(pheno.data.ili.v2)
```


## Strata frequency  
```{r eval=TRUE}
frequency.of.strata <- table(pheno.data.ili.v2$stratum); mean(frequency.of.strata); frequency.of.strata
```

## Data for testing
```{r}
# remove firt column from asv_dat it is subject id
asv.iliv2.dat <- dplyr::select(asv.iliv2, -ID) 
# Create "data.set":
data.set.ili.v2 <- cbind(pheno.data.ili.v2,asv.iliv2.dat)

row.names(data.set.ili.v2) <- NULL
# get only bacterial names

#ncols.pheno.v2 <- ncol(pheno.data.v2) +1
#nocl.v2 <- ncol(data.set.v2)-5

ncols.pheno.ili.v2 <- ncol(pheno.data.ili.v2) +1

nocl.ili.v2 <- ncol(data.set.ili.v2)-5
bacterial.abund.ili.v2 <- names(data.set.ili.v2)[ ncols.pheno.ili.v2:nocl.ili.v2]

other <- c("average","SD","prop.positive","entropy","CV")

#vars.test <- colnames(phenotypic.data.1)[c(2,5:39)]
vars.test.ili.v2 <- c(patho.vars)

data.for.testing.ili.v2 <- subset(data.set.ili.v2,
                              select=c(vars.test.ili.v2,
                                       bacterial.abund.ili.v2,
                                       other,"stratum"))
#head(data.for.testing)
#saveRDS(data.for.testing, "ms/taxa_test/ctrl_ili/data.for.testing.ctrl.ili.rds")
```

## Type of vars 
```{r eval=TRUE}

TYPES.ili.v2 <- c(rep("ordinal",nocl.ili.v2),
           rep("continuous",5),
           "categorical")

types.ili.v2 <- c(rep("categorical",length(vars.test.ili.v2)-2),
           rep("dirac.and.continuous",length(bacterial.abund.ili.v2)),
           rep("continuous",5),
           "categorical")
#colnames(data.for.testing)[1:10]

```


## Test associations   
```{r}

no.of.factors <- length(vars.test.ili.v2)
aux.columns <- (length(vars.test.ili.v2)+1):(length(colnames(data.for.testing.ili.v2))-6)
variables.of.interest <- names(data.for.testing.ili.v2)
#nominal.bound.on.FDR <- 0.20
log.scale <- TRUE

```


```{r eval=TRUE}
#dir.create("../../test_stat")
dir.create("06_taxa_associations/iliv2/")
set.seed(1210125)
check_association(data.for.testing.ili.v2,
                  B = 100000,
                  no.of.factors= no.of.factors,
                  aux.columns = aux.columns,
                  log.scale = TRUE,
                  nominal.bound.on.FDR = 0.20,
                  phen.data = pheno.data.ili.v2,
                  compare.label="ili_v2",
                  aux.TYPES=TYPES.ili.v2,
                  cat.types = types.ili.v2,
                  path_loc="06_taxa_associations/iliv2/",
                  name.of.stratification= "Sex and age-group")
```


```{r}
path.files <- "06_taxa_associations/iliv2/"
pattern.name <- "SELECTED.associations.between."
test.iliv2 <- combine_tables(path.files,pattern.name)
test.iliv2$Association <- ifelse(test.iliv2$sign == -1, "Negative", "Positive")
test.iliv2 <- test.iliv2 %>% 
    filter(p.value <= 0.05)

DT::datatable(test.iliv2)
```


# Check ILI V3 vs Pathogens  

```{r}

ps.iliv3 <- subset_samples(ps.main.2.genus, 
                         condition_status == "acute_ili_recovery")

# clean taxonomy
ps.iliv3 <- prune_taxa(taxa_sums(ps.iliv3) > 0, ps.iliv3)

```


## Chnage metadata

Add infection information from ili v1 to ili v3 microbiota 
```{r}

ili.v1.meta <- get_tibble(ps.iliv1, slot="sam_data")

ps.iliv3.sub <- subset_samples(ps.iliv3, participant_id %in% ili.v1.meta$participant_id)

ili.v1.meta <- ili.v1.meta %>% 
  filter(participant_id %in% sample_data(ps.iliv3.sub)$participant_id) %>% 
  as.data.frame() %>% 
  column_to_rownames("participant_id")

sample_names(ps.iliv3.sub) <- rownames(ili.v1.meta)
sample_data(ps.iliv3.sub) <- sample_data(ili.v1.meta)

table(sample_data(ps.iliv3.sub)$condition_status, sample_data(ps.iliv3.sub)$influenza_2014_any)
table(sample_data(ps.iliv3)$condition_status, sample_data(ps.iliv3)$haemophilus_2014_any)

ps.iliv3 <- ps.iliv3.sub
```


## Create taxa abundance table 
```{r}

asv.iliv3 <- transform_to_fractions(ps.iliv3, 
                                    det.thres = 0.001,
                                    prev.thres = 10/100)
dim(asv.iliv3)
#asv.tab.ctrl.ili[1:5,1:5]
```


## Prep metadata  

```{r}
#human_metapneumovirus_2014
#respiratory_syncytial_virus_2014
#human_parainfluenzavirus_2014_any 
#rhinovirus_2014


pheno.data.ili.v3 <- prep_metadata(ps.iliv3, 
                                   sam.ids = "Sample_Name", 
                                   ref.group="acute_ili", 
                                   compare= "condition_status", 
                                   confounder.vars = c("gender", "age_yrs_oct2014"),
                                   var.interest=all.vars)

head(pheno.data.ili.v3)


pheno.data.ili.v3$human_metapneumovirus_2014[pheno.data.ili.v3$human_metapneumovirus_2014==3] <- 2

pheno.data.ili.v3$human_metapneumovirus_2014[pheno.data.ili.v3$human_metapneumovirus_2014==3] <- 2
pheno.data.ili.v3$respiratory_syncytial_virus_2014[pheno.data.ili.v3$respiratory_syncytial_virus_2014==3] <- 2

pheno.data.ili.v3$human_parainfluenzavirus_2014_any[pheno.data.ili.v3$human_parainfluenzavirus_2014_any==3] <- 2

pheno.data.ili.v3$rhinovirus_2014[pheno.data.ili.v3$rhinovirus_2014==3] <- 2

#summary(pheno.data.ili.v2)
```


## Strata frequency  
```{r eval=TRUE}
frequency.of.strata <- table(pheno.data.ili.v3$stratum); mean(frequency.of.strata); frequency.of.strata
```

## Data for testing
```{r}
# remove firt column from asv_dat it is subject id
asv.iliv3.dat <- dplyr::select(asv.iliv3, -ID) 
# Create "data.set":
data.set.ili.v3 <- cbind(pheno.data.ili.v3,asv.iliv3.dat)

row.names(data.set.ili.v3) <- NULL
# get only bacterial names

#ncols.pheno.v2 <- ncol(pheno.data.v2) +1
#nocl.v2 <- ncol(data.set.v2)-5

ncols.pheno.ili.v3 <- ncol(pheno.data.ili.v3) +1

nocl.ili.v3 <- ncol(data.set.ili.v3)-5
bacterial.abund.ili.v3 <- names(data.set.ili.v3)[ ncols.pheno.ili.v3:nocl.ili.v3]

other <- c("average","SD","prop.positive","entropy","CV")

#vars.test <- colnames(phenotypic.data.1)[c(2,5:39)]
vars.test.ili.v3 <- c(patho.vars)

data.for.testing.ili.v3 <- subset(data.set.ili.v3,
                              select=c(vars.test.ili.v3,
                                       bacterial.abund.ili.v3,
                                       other,"stratum"))
#head(data.for.testing)
#saveRDS(data.for.testing, "ms/taxa_test/ctrl_ili/data.for.testing.ctrl.ili.rds")
```

## Type of vars 
```{r eval=TRUE}

TYPES.ili.v3 <- c(rep("ordinal",nocl.ili.v3),
           rep("continuous",5),
           "categorical")

types.ili.v3 <- c(rep("categorical",length(vars.test.ili.v3)-2),
           rep("dirac.and.continuous",length(bacterial.abund.ili.v3)),
           rep("continuous",5),
           "categorical")
#colnames(data.for.testing)[1:10]

```


## Test associations   
```{r}

no.of.factors <- length(vars.test.ili.v3)
aux.columns <- (length(vars.test.ili.v3)+1):(length(colnames(data.for.testing.ili.v3))-6)
variables.of.interest <- names(data.for.testing.ili.v3)
#nominal.bound.on.FDR <- 0.20
log.scale <- TRUE

```


```{r eval=TRUE}
#dir.create("../../test_stat")
dir.create("06_taxa_associations/iliv3/")
set.seed(1210125)
check_association(data.for.testing.ili.v3,
                  B = 100000,
                  no.of.factors= no.of.factors,
                  aux.columns = aux.columns,
                  log.scale = TRUE,
                  nominal.bound.on.FDR = 0.20,
                  phen.data = pheno.data.ili.v3,
                  compare.label="ili_v3",
                  aux.TYPES=TYPES.ili.v3,
                  cat.types = types.ili.v3,
                  path_loc="06_taxa_associations/iliv3/",
                  name.of.stratification= "Sex and age-group")
```

```{r}
path.files <- "06_taxa_associations/iliv3/"
pattern.name <- "SELECTED.associations.between."
test.iliv3 <- combine_tables(path.files,pattern.name)
test.iliv3$Association <- ifelse(test.iliv3$sign == -1, "Negative", "Positive")
test.iliv3 <- test.iliv3 %>% 
    filter(p.value <= 0.05)

DT::datatable(test.iliv3)
```


```{r}
sessionInfo()
```

